---
title: "Carrefour-Marketing-Project Applying Feature Selection"
author: "Dynasty"
date: "`r Sys.Date()`"
output:
  html_document: default
  word_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# **Introduction**

Carrefour has 13 outlets mostly located in the suburbs of Kenya's capital city, Nairobi. Their mission is to provide our customers with quality services, products and food accessible to all across all distribution channels.

## **Problem Statement**

The project aim to inform the marketing department on the most relevant marketing strategies that will result in the highest no. of sales (total price including tax).

## **Metrics Of Success**

1. Perform data cleaning and EDA.
2. Applying method of Feature selection ().
3. Providing insights gained from the analysis.
 
## **Data Understanding**

Libraries necessary
```{r}
library(data.table)
library(dplyr)
library(tidyverse)
library(ggplot2)
```


Loading the data set
```{r}
df <- read.csv("http://bit.ly/CarreFourDataset")
```

Previewing the data set
```{r}
head(df)
```

Number of records
```{r}
dim(df)
```
We have 1000 rows and 16 columns.


Checking the structure of the data set
```{r}
str(df)
```
We have 8 numeric columns, 5 categorical column and two columns with date/time.



## **Data Cleaning and Data Preparation**

**Checking for Missing Values**
```{r}
is.null(df)
```
We don't have Null values is the data set.

**Checking for Duplicates**
```{r}
duplicated_rows <- df[duplicated(df),]
duplicated_rows
```
We don't have duplicates in our data set


**Checking for Outliers**

Selecting the numeric columns
```{r}
num <- df[,c(6:8, 12, 14:16)]

num
```

```{r}
ggplot(num) +
  aes(x = "", y = Unit.price) +
  geom_boxplot(fill = "#0c4c8a") +
  theme_minimal()
```

```{r}
ggplot(num) +
  aes(x = "", y = Quantity) +
  geom_boxplot(fill = "#0c4c8a") +
  theme_minimal()
```


```{r}
ggplot(num) +
  aes(x = "", y = Tax) +
  geom_boxplot(fill = "#0c4c8a") +
  theme_minimal()
```


```{r}
ggplot(num) +
  aes(x = "", y = cogs) +
  geom_boxplot(fill = "#0c4c8a") +
  theme_minimal()
```


```{r}
ggplot(num) +
  aes(x = "", y = gross.income) +
  geom_boxplot(fill = "#0c4c8a") +
  theme_minimal()
```




```{r}
ggplot(num) +
  aes(x = "", y = Rating) +
  geom_boxplot(fill = "#0c4c8a") +
  theme_minimal()
```


```{r}
ggplot(num) +
  aes(x = "", y = Total) +
  geom_boxplot(fill = "#0c4c8a") +
  theme_minimal()
```




The box plots above clearly indicate we have a few outliers in Tax, cogs, gross.income and total column. However, they are important for our analysis so will keep them for now



# **Exploratory Data Analysis**

### **Univariate Analysis**

Previewing branch distribution
```{r}
branch <- (df$Branch)
branch.frequency <- table(branch)
branch.frequency
```


```{r}
barplot(branch.frequency,
  main="A bar chart showing Branch distribution",
  xlab="Branch",
  ylab = "Frequency",
  col=c("red","purple", "blue"),
  )
```



The bar graph above shows Branch A had the highest distribution with Brach c with the least even though the differences was very minimal.


Previewing Product line distribution
```{r}
pl <- (df$Product.line)
pl.frequency <- table(pl)
pl.frequency
```
```{r}
barplot(pl.frequency,
  main="A bar chart showing Product line distribution count",
  xlab="Product Line",
  ylab = "Count",
  col=c("#eb8060", "#b9e38d", "green", "yellow", "#a1e9f0", "#d9b1f0"),
  )
```

Most sales were from Fashion accessories product line while least sales were from Health and beauty product line. 


Previewing Customer type Distribution
```{r}
cust <- (df$Customer.type)
cust.frequency <- table(cust)
cust.frequency
```


```{r}
barplot(cust.frequency,
  main="A bar chart showing Customer type distribution count",
  xlab="Customer Type",
  ylab = "Count",
  col=c("red","purple"),
  )
```


Most customer were members customers. 

Previewing payment methods
```{r}
payment <- (df$Payment)
payment.frequency <- table(payment)
payment.frequency
```


```{r}
barplot(payment.frequency,
  main="A bar chart showing payment method distribution count",
  xlab="Payment Method",
  ylab = "Count",
  col=c("magenta","purple", "green"),
  )
```
Most customers use Ewallet and least payment method used by customers is via credit card.



Previewing Gender distribution
```{r}
gender <- (df$Gender)
gender.frequency <- table(gender)
gender.frequency
```


```{r}
barplot(gender.frequency,
  main="A bar chart showing Gender distribution count",
  xlab="Gender",
  ylab = "Count",
  col=c("magenta","purple"),
  )
```

Most customers were female


Previewing Unit prices distribution
```{r}
hist((df$Unit.price),  
main = "Histogram of Unit Price distribution",
     xlab = 'Unit Price', 
     ylab = 'Frequency',
     col = "blue")
```

Maximum unit prices were around 100.


```{r}
hist((df$Quantity),  
main = "Histogram of Quantity distribution",
     xlab = 'Quantity', 
     ylab = 'Frequency',
     col = "blue")
```


```{r}
hist((df$Tax),  
main = "Histogram of Tax distribution",
     xlab = 'Tax', 
     ylab = 'Frequency',
     col = "blue")
```


```{r}
hist((df$cogs),  
main = "Histogram of cogs distribution",
     xlab = 'cogs', 
     ylab = 'Frequency',
     col = "blue")
```


```{r}
hist((df$gross.income),  
main = "Histogram of gross income distribution",
     xlab = 'gross income', 
     ylab = 'Frequency',
     col = "blue")
```


```{r}
hist((df$Rating),  
main = "Histogram of Rating distribution",
     xlab = 'Rating', 
     ylab = 'Frequency',
     col = "blue")
```

```{r}
hist((df$Total),  
main = "Histogram of Total distribution",
     xlab = 'Total', 
     ylab = 'Frequency',
     col = "blue")
```



### **Bivariate Analysis**
```{r}
plot((df$Unit.price), (df$Tax), 
     main = "A scatterplot of Unit Price against Tax",
     xlab = 'Unit Price', 
     ylab = 'Tax')
```

The higher the unit price the higher the tax applied


```{r}
ggplot(df,aes(x=gross.income, y=Tax))+geom_line()+
ggtitle("A scatterplot of Gross income against Tax")+
theme(plot.title = element_text(hjust=0.5))
```

Tax increases with increase in Gross income.

```{r}
ggplot(df,aes(x=Date, y=Tax))+geom_line()
```


Tax has been increasing and decreasing over time



### **Multivariate Analysis**

getting a summary for the data set from our numeric columns
```{r}
summary(num)
```
We can observe the following:
a. The  maximum unit price was 99.96 and the minimum was 10.08. 
b. The maximum quantity ordered was 10 while the minimum was 1.
c. The highest Tax applied was 49.65 while the least was 0.5.
d. The highest rating was 10 while the least rating was 4.


Ploting a correlation plot
```{r}
library(corrplot)
corrplot(cor(num))
```


# **Feature Selection**
## **a. Filter Method**
```{r}
library(caret)
library(corrplot)
```
Will be using the numeric columns for this analysis, so let's preview
```{r}
head(num)
```

First will be calculating the the correlation matrix of our attributes
```{r}
correlationMatrix <- cor(num)
```

Find attributes that are highly correlated
```{r}
highlyCorrelated <- findCorrelation(correlationMatrix, cutoff=0.90)
```

```{r}
names(num[,highlyCorrelated])
```
We can see Cost of goods sold, Tax and total have a very high correlation.

Will remove these highly correlated features
```{r}
num2<-num[-highlyCorrelated]
num2
```
We are left with features that don't have high correlation

Plot these findings
```{r}
par(mfrow = c(1, 2))
corrplot(correlationMatrix, order = "hclust")
```
The graph shows all the seven features in our data set and it clearly shows correlated features


```{r}
corrplot(cor(num2), order = "hclust")
```
The above plot shows features with no correlation or if there is any correlation it's very minimal.


## **b. Wrapper Methods** 
```{r}
library(clustvarsel)
library(mclust)
```


Performing Sequential forward greedy search (default)

```{r}
out = clustvarsel(num)
out
```

The selection algorithm would indicate that the subset we use for the clustering model is composed of variables: Quantity, Total, Unit.price and Rating while the other variables should be rejected. 
Having identified the variables that we use, we proceed to build the clustering model
```{r}
Subset1 =num[,out$subset]
mod = Mclust(Subset1)
summary(mod)
```


```{r}
plot(mod,c("classification"))
```

The graph shows that the optimal subset of variables are Quantity, Unit.price, Rating and Total.

Using wrapper method are almost similar in their feature selection with filter method except for total and gross income

## **c. Embedded Methods**
This is a weighted subspace clustering algorithm 

Loading necessary libraries
```{r}
library(wskm)
library("cluster")
```

Creating our model
```{r}
set.seed(2)
model <- ewkm(num, 3, lambda=2, maxiter=1000)
```

Cluster Plot against our principal components
```{r}
clusplot(num, model$cluster, color=TRUE, shade=TRUE,
         labels=2, lines=1,main='Cluster Analysis for Sales features')
```

Checking for weights stored in the model
```{r}
round(model$weights*100,2)
```

Weights  measure of the relative importance of each variable with regards to the membership of the observations to that cluster. 
The table above shows the features of importance includes Rating, gross income and Tax.


In hindsight all the three methods features of importance are:
i. Filter method: Quantity, Unit.price, Rating and gross.income.
ii. Wrapper method: Quantity, Unit.price, Rating and Total.
iii. Embedded method: Rating, gross income and Tax


**Conclusion**
* We can confidently advise Carrefour marketing department on coming up with the 
most relevant marketing strategies they should consider unit.price, Rating, Quantity, Tax
and gross.Income



